A user-extensible and adaptable parser architecture

### 2008-08-15 doc/logparser-implementation.tex
Next step in the great renaming - renaming of graphs. In the process I removed some generated gnuplot input files from the repository.
Renamed actions because Carl can't understand that the IGNORE action doesn't mean that the log line wasn't parsed, so other people might have the same problem:     IGNORE              => UNINTERESTING     REJECTION           => DELIVERY_REJECTED     SAVE_BY_QUEUEID     => SAVE_DATA

Every interesting\footnote{All log lines are interesting unless the action
of the rule which parses them is \texttt{INFO} or \texttt{IGNORE}, as
described in \sectionref{actions-in-detail}.} log line will have an entry
in the results table, e.g.\ rejecting an \SMTP{} command, delivering a
mail, or bouncing a mail (see \sectionref{postfix_action} for typical
values of \texttt{postfix\_action}).  Each row is associated with a single
connection, though there may be many results per connection.

@@

@@

    \item [IGNORE] This rule just returns successfully; it is used when a
        line needs to be parsed for completeness but does not either
        provide any useful data or require anything to be done.

@@

    \item [REJECTION] Deal with Postfix rejecting an \SMTP{} command from
        the remote client: log the rejection with a mail if there is a
        queueid in the log line, or with the connection if not.

    \item [SAVE\_BY\_QUEUEID] Find the correct mail based on the queueid in
        the log line, and save the data extracted by the \regex{} to it.

@@

@@

\showgraph{build/plot-action-distribution}{Distribution of rules per
action}{Distribution of rules per action}

>>>>

Every interesting\footnote{All log lines are interesting unless the
postfix\_action of the rule which parses them is \texttt{INFO} or
\texttt{IGNORE}, as described in \sectionref{actions-in-detail}.} log line
will have an entry in the results table, e.g.\ rejecting an \SMTP{}
command, delivering a mail, or bouncing a mail (see
\sectionref{postfix_action} for typical values of
\texttt{postfix\_action}).  Each row is associated with a single
connection, though there may be many results per connection.

@@

    \item [DELIVERY\_REJECTED] Deals with Postfix rejecting an \SMTP{}
        command from the remote client: log the rejection with a mail if
        there is a queueid in the log line, or with the connection if not.

@@

@@

    \item [SAVE\_DATA] Find the correct mail based on the queueid in
        the log line, and save the data extracted by the \regex{} to it.

@@

    \item [UNINTERESTING] This rule just returns successfully; it is used when a
        line needs to be parsed for completeness but does not either
        provide any useful data or require anything to be done.

@@

\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action}

### 2008-08-15 doc/sgai-2008.tex
Renamed actions because Carl can't understand that the IGNORE action doesn't mean that the log line wasn't parsed, so other people might have the same problem:     IGNORE              => UNINTERESTING     REJECTION           => DELIVERY_REJECTED     SAVE_BY_QUEUEID     => SAVE_DATA

Each action contains the code required to deal with a particular type of
log lines, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between log lines.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{REJECTION}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_BY\_QUEUEID}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{IGNORE}: the \texttt{IGNORE} action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects
(neither \texttt{REJECTION} nor \texttt{IGNORE} imply that the input is
ungrammatical or unparsed).  The framework warns about every unparsed log
line, to alert the user that they need to alter or extend their ruleset;
the Postfix log parser successfully parses every input line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
remaining actions have only one or two associated rules: in some cases this
is because there will only ever be one type of log line for that action,
e.g.\ all log lines showing that a remote client has connected are matched
by a single rule and handled by the \texttt{CONNECT} action; in other cases
the action is required to address a deficiency in the log files, or a
complication which arises during parsing.  Taking the \texttt{CONNECT}
action as an example, it creates a new data structure for each client
connecting, and saves the data extracted by the rule (in memory,
subsequently to be entered into the database).  If the data structure
already exists it is treated as a symptom of a bug, and the action issues a
warning containing the full contents of the data structure, plus the log
line which has just been parsed.

>>>>

Each action contains the code required to deal with a particular type of
log lines, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  The distribution is skewed
for the same reason the architecture is so successful in enabling users to
parse new log lines: most new log lines can be parsed by a combination of a
new rule and an existing action.  It is expected that parsers utilising
this architecture will have a high ratio of rules to actions, hopefully
resulting in simpler rules and clearer distinctions between log lines.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the action
which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines (such as rejections), but that do not by themselves represent
interesting events.  The third most common action is, perhaps surprisingly,
\texttt{UNINTERESTING}: this action does nothing when executed,
allowing uninteresting log lines to be parsed without causing any effects
(it does not imply that the input is
ungrammatical or unparsed).  The framework warns about every unparsed log
line, to alert the user that they need to alter or extend their ruleset;
the Postfix log parser successfully parses every input line in the
\numberOFlogFILESall{} log files it is currently being tested with.  The
remaining actions have only one or two associated rules: in some cases this
is because there will only ever be one type of log line for that action,
e.g.\ all log lines showing that a remote client has connected are matched
by a single rule and handled by the \texttt{CONNECT} action; in other cases
the action is required to address a deficiency in the log files, or a
complication which arises during parsing.  Taking the \texttt{CONNECT}
action as an example, it creates a new data structure for each client
connecting, and saves the data extracted by the rule (in memory,
subsequently to be entered into the database).  If the data structure
already exists it is treated as a symptom of a bug, and the action issues a
warning containing the full contents of the data structure, plus the log
line which has just been parsed.
